---
title: "Linear Regression Tutorial"
date: 2024-06-20
output: html_document
---

# Introduction

In these exercises, we will be modeling the prices at which homes were sold in King County, Washington (Seattle, Tacoma, etc) between May 2014 and May 2015. This data set was downloaded from [Kaggle](https://www.kaggle.com/datasets/harlfoxem/housesalesprediction), and is in the file `kc_house_data.csv`. Run the code block below to load this data into a data.frame called `df_home`:

```{r}
df_home <- read.csv("../data/kc_house_data.csv")
```

Our goal in this exercise is to evaluate how different features of a home influence the `price` for which it sells. With our outcome variable being a numeric quantity, we 

Before diving into the meat of this tutorial, it might be helpful to go over a few functions in R that will let you get acquainted with the contents of a data.frame:

-   `head`: print the first 6 rows of a data.frame
-   `summary`: generate a summary of each variable in the data.frame
-   `str`: another way to get a quick summary of what kinds of data are in the data.frame

Note: you can pull up documentation by typing `?` followed by the name of the function (e.g.: `?str`) in the R console to pull up documentation for the function. Try this with the head function, and tell me: which paramter would we change to show more or less than the first 6 rows of a data.frame?

Run the following cell to get a quick overview of the types of data in our data.frame:

```{r}
summary(df_home)
```

As a last bit of data prep before we dive into linear regression, we're going to re-code the `waterfront` variable as a factor: it's a *dummy variable* which indicates whether or not the property is on a waterfront (0 = no, 1 = yes):

```{r}
df_home$waterfront <- factor(df_home$waterfront, levels = c(0,1), labels=c("N","Y"))

summary(df_home$waterfront)
```

We'll also re-code the `zipcode` variable as a categorical variable, since it does not make sense to treat this variable as a numeric value:

```{r}
df_home$zipcode <- factor(df_home$zipcode)

summary(df_home$zipcode)
```

Note: there are other variables that could be considered categorical (and in some cases, ordinal) in this data set as well. For now, we'll keep these in the data.frame as numeric values; later, we'll see the consequences of this choice.

## Data Description

Stolen from this [kaggle post](https://www.kaggle.com/datasets/harlfoxem/housesalesprediction/discussion/207885), here are definitions for all our variables:

* id - Unique ID for each home sold 
* date - Date of the home sale 
* price - Price of each home sold 
* bedrooms - Number of bedrooms 
* bathrooms - Number of bathrooms, where .5 accounts for a room with a toilet but no shower  
* sqft_living - Square footage of the interior living space 
* sqft_lot - Square footage of the land space 
* floors - Number of floors 
* waterfront - A dummy variable for whether the apartment was overlooking the waterfront or not 
* view - An index from 0 to 4 of how good the view of the property was 
* condition - An index from 1 to 5 on the condition of the apartment
* grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design
* sqft_above - The square footage of the interior housing space that is above ground level 
* sqft_basement - The square footage of the interior housing space that is below ground level
* yr_built - The year the house was initially built 
* yr_renovated - The year of the house’s last renovation
* zipcode - What zipcode area the house is in
* lat - Lattitude 
* long - Longitude 
* sqft_living15 - The square footage of interior housing living space for the nearest 15 neighbors
* sqft_lot15 - The square footage of the land lots of the nearest 15 neighbors

# Simple Linear Regression

In the following exercises we'll be exploring the influence of individual explanatory variables on the sale price of a home.

## Numeric predictor variables

We will first use linear regression to assess the relationship between the size of the house's interior living space (``)

Before fitting any models, let's make some scatter plots to visually inspect the relationship between these predictors and sale price.

### Exercise 1

Create scatter plots with price on the Y axis, and `sqft_living` on the X-axis

```{r}
# replace ??? with parameters for the plot function to generate scatter plot of price and living space
# need help? try typing "?plot" (without quotes) into the R console
plot(???)
```

Question: in general, does house price tend to increase with the size of living space?
Answer: 

### Exercise 2

Using linear regression, evaluate the relationship between `price` and `sqft_living`

```{r}
model1 <- lm(???, df_home) # replace ??? with a formula to specify our model: price = a + b * sqft_living
                           # reminder: the format of a formula for simple linear regression is outcome ~ predictor
```

Print out the summary of the model

```{r}
???(model1) # replace ??? with the name of the function we use to get a summary of our model results
```

Q: What is our fitted model:
A: price = ??? + ??? x sqft_living

Q: Does the slope have a meaningful interpretation here?
Hint: When would a house be predicted to have a price of -43580.743? Does this make sense?
A: 

Q: Interpret the coefficient for sqft_living. Is a higher sqft_living value associated with a higher price?
A:

### Exercise 3

In this exercise, we'll look at how linear regression models can also be used to evaluate the relationship between a numeric outcome and a categorical predictor. In particular, we will evaluate whether or not the mean sales price of a waterfront property is different from non-waterfront properties.

In your basic stats review lecture, you covered how the t test could be used to compare the means of two groups. We can very much do the same thing for our comparison of the waterfront and non-waterfront properties:

```{r}
t.test(price ~ waterfront, df_home) # notice how we can use the "formula, data.frame" parameterization for this function
```

Based on the above t-test, we would conclude that the mean price for waterfront properties is significantly different from the mean price of non-waterfront properties. Would we reach the same conclusion if we used a linear model instead?

First, make a boxplot of prices for both waterfront and non-waterfront properties:

```{r}
boxplot(price ~ waterfront, df_home) # replace ???s to get a boxplot of price values stratified by waterfront
                             # hint: formula follows same outcome ~ predictor pattern
```

Q: Visually, does it look like there is a difference in the distributions of prices between waterfront (1) and non-waterfront (0) properties?
A: 
